---
title: "R Notebook"
output: html_notebook
---

```{r, include=FALSE}
library(tidyverse)
library(lme4)
library(lmerTest)
library(knitr)
library(ggplot2)
library(plotrix)
library(kableExtra)
library(gridExtra)
```

### Question 1: Group Differences?

```{r}
get_group_differences <- function(f, to_be_table, measure) {
  data <- read.delim(f,sep=",")
  # Extract information from file name
  stripped = strsplit(f,'_')[[1]]
  word = stripped[length(stripped)-1]
  sentence = strsplit(stripped[length(stripped)],'.c')[[1]][1]

  # Get mean and standard deviation by group (chr vs hc)
  data <- data %>% filter(data[[measure]] < 998.0)
  data <- reshape(data, idvar=c("Participant", "Time", "Group"), timevar = "Question", direction="wide")
  
  if (measure == 'Coherence') {
    data$measure.avg <- rowMeans(data[,c("Coherence.challenge", "Coherence.selfdefining", "Coherence.turningpoint", "Coherence.unusual")], na.rm=TRUE)
  } else {
    data$measure.avg <- rowMeans(data[,c("Derailment.challenge", "Derailment.selfdefining", "Derailment.turningpoint", "Derailment.unusual")], na.rm=TRUE)
  }
  chr <- subset(data, Group == 'chr')
  hc <- subset(data, Group == 'hc')

  tt <- t.test(chr$measure.avg, hc$measure.avg, alternative = "less")
  tstat <- tt$statistic
  pvalue <- tt$p.value
  
  # Add all information to table and return 
  to_be_table[nrow(to_be_table)+1,] = c(word, sentence, round(mean(chr$measure.avg), digits = 3), round(mean(hc$measure.avg), digits = 3), round(sd(chr$measure.avg), digits = 2), round(sd(hc$measure.avg), digits = 2), round(tstat, digits = 2), round(pvalue, digits = 3))
  return(to_be_table)
}
```


```{r}
file_path = "/Users/khit/Box Sync/Postdoc/Research/narratives_clean/results/nar_mainqonlynofillers"
files <- list.files(path=file_path, pattern="*.csv", full.names=TRUE, recursive=FALSE)
# print(files)
# setwd(file_path)

basetable <- setNames(data.frame(matrix(ncol = 8, nrow = 0)), c("word", "sentence", "CHR mean", "HC mean", "CHR std", "HC std", "t-statistic", "p-value"))

# Get coherence table
coh <- lapply(files, get_group_differences, to_be_table = basetable, measure = 'Coherence')
coherence_table <- do.call(rbind, coh)
kable(coherence_table, caption = "Coherence = on average how similar are neighboring sentences")

# Get derailment table (referred to as "tangentiality" in the paper)
der <- lapply(files, get_group_differences, to_be_table = basetable, measure = 'Derailment')
derailment_table <- do.call(rbind, der)
kable(derailment_table, caption = "Derailment = slope of progressive coherence scores")
```


```{r}
# Read in one file
data <- read.delim("/Users/khit/Box Sync/Postdoc/Research/narratives_clean/results/nar_mainqonlynofillers/nar_mainqonlynofillers_glove_pos.csv",sep=",")
partinfo <- read.delim("/Users/khit/Box Sync/Postdoc/Research/narratives_clean/symps+demo/clpsych2021_part_info.csv",sep=",")
names(data)[names(data) == 'Coherence'] <- 'coh'
data$Derailment <- NULL
data$Tangentiality <- NULL
data$Iteration <- NULL
data <- reshape(data, idvar=c("Participant", "Time", "Group"), timevar = "Question", direction="wide")
data$coh.avg <- rowMeans(data[,c("coh.challenge", "coh.selfdefining", "coh.turningpoint", "coh.unusual")], na.rm=TRUE)
data <- merge(data, partinfo, by=c("Participant", "Time", "Group"))
```

```{r}
# Plot group differences CHR vs. HC
p <- ggplot(data,aes(x=coh.avg,color=Group)) + geom_density()
p <- ggplot(data, aes(x=Group, y=coh.avg, color=Group)) + geom_violin() + geom_jitter(shape=16, position=position_jitter(0.2)) + ylab("Mean Coherence")
p
```

### Question 2: Correlations with Symptoms?

```{r, include = FALSE}
chrdata <- subset(data, Group == 'chr')

# p5corr
p5corr <- ggplot(chrdata, aes(x=sips.p5, y=coh.avg)) + geom_point() + xlab("SIPS P5: Disorganized Communication") + ylab("Mean Coherence") + geom_smooth(method = "lm", se = TRUE)
# p5corr
round(cor.test(chrdata$sips.p5, chrdata$coh.avg)$estimate, digits=3)
round(cor.test(chrdata$sips.p5, chrdata$coh.avg)$p.value, digits=3)

# n5 corr
n5corr <- ggplot(chrdata, aes(x=sips.n5, y=coh.avg)) + geom_point() + xlab("SIPS N5: Ideational Richness") + ylab("Mean Coherence") + geom_smooth(method = "lm", se = TRUE)
round(cor.test(chrdata$sips.n5, chrdata$coh.avg)$estimate, digits=3)
round(cor.test(chrdata$sips.n5, chrdata$coh.avg)$p.value, digits=3)

# d2 corr
d2corr <- ggplot(subset(chrdata, sips.d2 < 998.0), aes(x=sips.d2, y=coh.avg)) + geom_point() + xlab("SIPS D2: Bizarre Thinking") + ylab("Mean Coherence") + geom_smooth(method = "lm", se = TRUE)
round(cor.test(chrdata$sips.d2, chrdata$coh.avg)$estimate, digits=3)
round(cor.test(chrdata$sips.d2, chrdata$coh.avg)$p.value, digits=3)

# ptotal corr
ptotalcorr <- ggplot(chrdata, aes(x=sips.ptotal, y=coh.avg)) + geom_point() + xlab("SIPS Positive Total") + ylab("Mean Coherence") + geom_smooth(method = "lm", se = TRUE)
round(cor.test(chrdata$sips.ptotal, chrdata$coh.avg)$estimate, digits=3)
round(cor.test(chrdata$sips.ptotal, chrdata$coh.avg)$p.value, digits=3)

# nottal corr
ntotalcorr <- ggplot(chrdata, aes(x=sips.ntotal, y=coh.avg)) + geom_point() + xlab("SIPS Negative Total") + ylab("Mean Coherence") + geom_smooth(method = "lm", se = TRUE)
round(cor.test(chrdata$sips.ntotal, chrdata$coh.avg)$estimate, digits=3)
round(cor.test(chrdata$sips.ntotal, chrdata$coh.avg)$p.value, digits=3)

# dtotal corr
dtotalcorr <- ggplot(subset(chrdata, sips.d2 < 998.0), aes(x=sips.dtotal, y=coh.avg)) + geom_point() + xlab("SIPS Disorganized Total") + ylab("Mean Coherence") + geom_smooth(method = "lm", se = TRUE)
round(cor.test(chrdata$sips.dtotal, chrdata$coh.avg)$estimate, digits=3)
round(cor.test(chrdata$sips.dtotal, chrdata$coh.avg)$p.value, digits=3)

grid.arrange(p5corr, ptotalcorr, n5corr, ntotalcorr, d2corr, dtotalcorr, nrow = 3)

```

### Question 3a: Correlations with Surface-Level Linguistic Factors

```{r}
# Sentence Length - sig correlation
round(cor.test(data$wordspersent.avg, data$coh.avg)$estimate, digits=3)
round(cor.test(data$wordspersent.avg, data$coh.avg)$p.value, digits=3)
wordspersentcorr <- ggplot(data, aes(x=wordspersent.avg, y=coh.avg)) + geom_point() + xlab("Average Number of Words Per Sentence") + ylab("Mean Coherence") + geom_smooth(method = "lm", se = TRUE)

rval <- round(cor.test(data$wordspersent.avg, data$coh.avg)$estimate, digits=3)
pval <- round(cor.test(data$wordspersent.avg, data$coh.avg)$p.value, digits=10)
# Note: I've hard-coded p < 0.001 here
title <- paste0('R=', rval, ' (p<0.001)')
ggplot(data, aes(x=wordspersent.avg, y=coh.avg, color=Group)) +
  geom_point(size=3) + 
  geom_smooth(method = "lm", se = TRUE, aes(group=1), color = "Black") +
  labs(
    x = "Average Sentence Length\n(in words)",
    y = "Mean Coherence") +
  ggtitle(title) +
  theme(plot.title = element_text(hjust = 0.5))

# MATTR - no correlation
round(cor.test(data$MATTR.avg, data$coh.avg)$estimate, digits=3)
round(cor.test(data$MATTR.avg, data$coh.avg)$p.value, digits=3)
mattrcorr <- ggplot(data, aes(x=MATTR.avg, y=coh.avg)) + geom_point() + xlab("Lexical Diversity\n(Moving Average Type-to-Token Ratio)") + ylab("Mean Coherence") + geom_smooth(method = "lm", se = TRUE)

rval <- round(cor.test(data$MATTR.avg, data$coh.avg)$estimate, digits=3)
pval <- round(cor.test(data$MATTR.avg, data$coh.avg)$p.value, digits=3)
title <- paste0('R=', rval, ' (p=', pval, ')')
ggplot(data, aes(x=MATTR.avg, y=coh.avg, color=Group)) +
  geom_point(size=3) + 
  geom_smooth(method = "lm", se = TRUE, aes(group=1), color = "Black") +
  labs(
    x = "Lexical Diversity\n(Moving Average Type-to-Token Ratio)",
    y = "Mean Coherence") +
    ggtitle(title) +
    theme(plot.title = element_text(hjust = 0.5))

```

### Question 3a: Randomizing Words Simulations

```{r, echo = FALSE, include = FALSE}
# Read in data to get Monte Carlo difference distributions
det <- read.delim("/Users/khit/Box Sync/Postdoc/Research/narratives_clean/results/nar_mainqonlywithfillers/nar_mainqonlywithfillers_glove_pos.csv",sep=",")
mc <- read.delim("/Users/khit/Box Sync/Postdoc/Research/narratives_clean/results/nar_mainqonlywithfillers/nar_mainqonlywithfillers_glove_pos_randomwords.csv",sep=",")

det <- det %>% group_by(Participant,Group) %>% summarize(meanCoh = mean(Coherence))
det <- det %>% group_by(Group) %>% summarize(Coherence=mean(meanCoh))
det <- spread(det, Group, Coherence)
det$Diff <- det$hc - det$chr

x <- mc %>% group_by(Participant,Group,Iteration) %>% summarize(meanCoh = mean(Coherence))
x <- x %>% group_by(Iteration,Group) %>% summarize(Coherence=mean(meanCoh))
x <- spread(x, Group, Coherence)
x$Diff <- x$hc - x$chr
mc <- x
mc <- mc[order(-mc$Diff),]
# kable(mc)
p <- ggplot(mc, aes(x=Diff)) + 
  geom_density() +
  geom_vline(xintercept = det$Diff, colour="blue", linetype = "longdash", size = 1) +
  xlab("HC-CHR Difference in Coherence") + ylab("Density")

```

### Question 3b: Correlations with Sociodemographic Factors (Continuous)

```{r}
# Function to calculate correlation
get_correlation <- function(data, xvar, yvar, xlabel, ylabel, showgroups) {
  rval <- round(cor.test(data[[xvar]], data[[yvar]])$estimate, digits=3)
  pval <- round(cor.test(data[[xvar]], data[[yvar]])$p.value, digits=3)
  title <- paste0('R=', rval, ' (p=', pval, ')')
  if (showgroups == 1) {
    corrplot <- ggplot(data, aes(x=data[[xvar]], y=data[[yvar]], color=Group)) + geom_point(size=2) + xlab(xlabel) + ylab("Mean Coherence") + geom_smooth(method = "lm", se = TRUE, aes(group=1), color = "Black") + ggtitle(title) + theme(plot.title = element_text(hjust = 0.5)) 
  } else {
    corrplot <- ggplot(data, aes(x=data[[xvar]], y=data[[yvar]])) + geom_point() + xlab(xlabel) + ylab("Mean Coherence") + geom_smooth(method = "lm", se = TRUE) + ggtitle(title) + theme(plot.title = element_text(hjust = 0.5))
  }
  
  return(corrplot)
}

# Calculate correlations with continuous variables
agecorr <- get_correlation(data, 'age', 'coh.avg', 'Age', 'Mean Coherence', 0)
yearseducorr <- get_correlation(data, 'years_edu', 'coh.avg', 'Education Level', 'Mean Coherence', 0)
wratcorr <- get_correlation(data, 'WRAT', 'coh.avg', 'WRAT', 'Mean Coherence', 1)
```

### Question 3b: Correlations with Sociodemographic Factors (Discrete)

```{r}

# Racial Identity
data_race <- subset(data, race %in% c(5,8))
data_race$race[data_race$race==5] <- 2
data_race$race[data_race$race==8] <- 1

data_race$race <- as.factor(data_race$race)
data_race$Group <- as.factor(data_race$Group)
p <- ggplot(data_race, aes(x=race, y=coh.avg, fill=Group, na.rm = TRUE)) +
  geom_boxplot(alpha = .7, outlier.shape = NA) +
  scale_fill_discrete(name = "Group", labels = c("CHR", "HC")) +
  geom_point(position=position_jitterdodge()) +
  stat_summary(fun.y="mean", geom="point", size=3,
    position=position_dodge(width=0.75), color="white", show_guide=FALSE) +
  scale_x_discrete(labels=c("White","Black")) +
  labs(
    x = "Race",
    y = "Coherence"
  )

```

### Question 3b: Regression - Sentence Length, WRAT, Racial Identiy

```{r}
data$GroupCode <- ifelse(data$Group=="hc",-.5, .5)
data$RaceCode <- ifelse(data$race=='1',-.5,.5)

data.lmer <- lm(coh.avg~WRAT+wordspersent.avg+RaceCode, data=data,
                  control = lmerControl(optimizer = "bobyqa",optCtrl=list(maxfun=100000)),verbose = 2)

summary(data.lmer)
```



